Going from coarse landings data to fine scale species distribution

### 2022-06-02 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

**Attention:** L'introduction n'est pas finie

@@

Reconciling spatial (and eventually temporal) scales properly when different set of observations do not have the same resolution is a well known issue in geography, ecology, agriculture, geology and statistics [@gotway2002combining]. Inferring values of a variable at points or blocks different from those that have been observed is commonly refered as the *change of support problem* (COS) in the statistical literature [@gotway2002combining and @gelfand2010misaligned]. COS issues can be classified within 2 categories - that are not necessarily mutually exclusive - [@pacifici2019resolving]:

@@

- (2) 'modifiable areal unit' problem (MAUP) when data are aggregated over increasingly larger geographic scales (e.g. collected at point level but regrouped/declared at coarse level - see for instance @gilbert2021integrating). Our case of application clearly refers to MAUP. Note that within this category, another category exist and is commonly refered to the 'ecological fallacy'. This occures when the individual response to a covariate differs from the response estimated from grouping individuals. This means that conclusions from a fine-resolution analysis can differ from an analysis at a coarser scale based on the aggregation of the fine-resolution data.

@@

#  Introduction 
* Commencer par la motivation peche
--> knowledge of essential habitat --> Species distribution model --> link between biomass and habitat to identify essential habitat
-->  Data from scientific survey, precisely geolocalized standard sampling protocol, expensive and time consuming
--> On the other hand declaration data, spatialized at coarser level  but cheap and reglementaire
Combining the two sources of data.

This problem is not unique, other exampl in terms of apllication

Change of support form fine to coarse sptial level

A solution the reallocation process

** Pour la discu, simulation conditionnelle ? **

We proposed an alternative approach to account for the change of support and we compare it to the approach proposed in alglave but also to a golden standard to assess the consequences of these methods.

Paper organized as follow

Section 2. presentation of the reference model and detail the reallocation process

Section 3. Presnetation of an alternaion way to deal with change of support

Section 4. simulation study to asses the bias and/or the loss of precision of the two methodes compared with the golden standard

Section 5. Appllication on sole of the Bay of Biscay

Section 6. Discu/conclu

@@

In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen. This section provides a brief overview of the key ingredient of this model and raises some concerns on the spatialization of the commercial catch data.

Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ a spatial random field which represents the biomass for a species of interest.  $\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$ and $\Gamma$ a field of covariate.  As the catch are   zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ with a mixture of a Dirac mass at $0$ and a Log Normal distribution, the proportion of the mixture being defined  conditionally on the random field $\bm{S}$.

\begin{equation}
\label{eq:defY}
Y_i | S(x_i)  \overset{ind}{\sim} \My{p_i}{\mu_i}{\sigma^2},
\end{equation}
with $p_i \defeq \exp(-e^\xi S(x_{i}))$ the proportion of the mixture, $\mu_i\defeq 
\frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ a transformation of its variance variance. This parametrization (see SMXXXX) of the mixture model corresponds to:
\begin{align}
\P\left (Y_i =0 | S(x_i) \right) & = p_i=  \exp(-e^\xi S(x_{i}) ) \\  \nonumber
\E\left ( Y_i \vert Y_i >0 , S(x_i)\right)  & =  \mu_i = \frac{S(x_i)}{1-p_i},\\  \nonumber
\V ar( Y_i\vert Y_i >0, S(x_i) )& = \mu_i^2 (e^{\sigma^2}-1), \cr \nonumber
\end{align}

<!-- 
Je pense qu'il ne faut pas parler de la grille discrete tout de suite car S est défini sur un espace continu et moralement l'intensité du processus de poisson aussi. le passage à la grille est la pour simplifier l'expression de la vraisemblance, on peut dire qu'on fait une approximation pour des raisons numériques

We consider a discrete study domain where $x \in [\![1,n]\!]$ stands for the discrete locations. $\beta . \Gamma(x)$ is the covariate term with $\Gamma$ the matrix of design of the covariates and $\beta$ the effect of the covariates. $\delta(x)$ is a GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$). -->

@@

<!-- however those fishinf locations are actually derived from a different dataset containing the hourly localisation of the the commercial vessels, the VMS dataset. This dataset is processed to identify the fishing location among all the recorded locations. -->

<!-- At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch). Observations are supposed to follow some distribution $\mathcal{L}_Y$. -->

<!-- $$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma^2)$$ -->

<!-- In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022. $\mathcal{L}_Y$ is decomposed it 2 parts: -->

<!-- - the probability to obtain a zero-catch which is modelled as a Bernoulli variable with probability $p_{i}=\exp(-e^\xi .S(x_{i}))$. $\xi$ is a parameter controlling zero-inflation. When $\xi$ increases, the amount of zero in the data decreases. -->

<!-- - if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a positive continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$. -->

<!-- $$ -->
<!-- \operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) = -->
<!-- \left\{ -->
<!--     \begin{array}{ll} -->
<!--         p_i & \text { if } y_{i}=0 \\ -->
<!--         \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0 -->
<!--     \end{array} -->
<!-- \right. -->
<!-- $$ -->

@@

$$D=\sum_{x_i \in \mathcal{A}_D}{Y(x_i)}.$$

Since the original model proposed in @alglave_combining_2022 is defined at the individual catch scale, those catch are  derived from the logbooks data combined with the VMS data. This is classically done in fisheries science REFREFEREFREF by using a uniform reallocation process. 
This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_{D_j}$ associated with declaration $D_j$ and define for each $x_i\in\mathcal{A}_{D_j},$ the associated reallocated individual catch $\Yr(x_i)\defeq D_j/m_j$. As a direct consequence, the so reconstructed individual catch are similar to quantity widely used in Fishereis Sceince named  Catch Per Unit of Effort (CPUE), effort being measure in hours. As noted by @alglave_combining_2022, this process has several drawbacks.
First, as a consequence of the reallocation process, the reconstructed individual catches tend to exhibit smoother pattern than the original catches. 
Second, the actual number of data is the total number of catch declarations while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations.
From a statistical point of view, this overestismation of the number of informative data tends to produce excessively narrow confidence intervals.

@@

<!-- In standard processing, $D_j$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$. This is what we call uniform (or proportional) reallocation. In this case, inference of species distribution is directly computed through reallocated $Y^*_i$ assuming these are the exact punctual observations. This strongly simplifies the actual process of observation and have most likely repercussion on model performance. -->

<!-- To overcome such limitations, an alternative is to consider that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables). Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->

To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the reconstructed individual catch $\Yr(x_i)$. As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables following a mixture distribution. This distribution has no  known analytical form. However the catch declaration data also exhibit some zero inflation and a long tail repartition of the values: a mixture model would be a good candidate. 
Therefore, we propose to use the same family of distribution for  $D_j$ than the one use for the individual catch $Y$: a mixture between a Dirac mass in $0$ and a lognormal distribution. This is not exact, as the distribution of $D_j$ should be defined as convolution of mixtures which  would result in an untractable model. We somehow approximate the sum of  the original mixture distributions  by a mixture distribution of the same kind and we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original proposition:

<!-- that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables). Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->

$$D_j | \mathcal{P}_j,\bm{S} \sim \Md{p^D_j}{\mu^D_j}{\sigma^{2,D}}$$

with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu^D_j$ the expected positive biomass, $p_j^D$ the proportion of the mixture and $\sigma^{2,D}$ the variance parameter.

<!-- As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$. Our approach is to match the moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_j$. This is what we call model-based reallocation. This model will be referred as the Declaration Model. In contrast, the model fitted on reallocated observations $Y^*_i$ will be referred as the $Y^*_i$ model. -->

<!-- In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | \mathcal{P}_j,S_{\mathcal{P}_j}$. -->

@@

In order to define compatible models for $Y$ the actual individual catches and $D$ the catch declarations, we choose to match the key quantities of the two distributions.

(1) As the $Y_1, \ldots, Y_{m_j}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $P(D_j = 0\vert \bm{S}, \mathcal{P}_j )$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.

\begin{align*}
\P(D_j = 0\vert \bm{S}, \mathcal{P}_j)  & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = p_j^D
\end{align*}

@@

$$\E(D_j \vert D_j > 0, \bm{S}, \mathcal{P}_j)= \frac{\sum_{x_i \in \mathcal{P}_j} \E(Y_i \vert S(x_i)) }{1-p_j^D} = \frac{\sum_{x_i \in \mathcal{P}_j} S(x_{i})}{1-p_j^D},$$
$$\V ar(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} \V ar(Y_{i})}{1-p_j^D} - \frac{\pi_j}{(1-p_j^D)^2}E(D_j)^2,$$
$$\mbox{with } \V ar(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i})).$$

By defining the observation process at the declaration level,  we expect to  avoid some of the drawbacks of the estimation based on reallocated individual catch data $Y^r$.

 The inference is based on  maximum likelihood approach with two simplifications. We use the Stochastic Partial Differential Equations (SPDE) approach to represent the spatial field and we use integrated Nested Laplace approximation to compute le likelihood. The stochastic random field is also approximated by a piecewise constant process defined on a fine grid. The optimization of the likelihood relies Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].

<!-- XXXX  Il faut aussi présenter la partie scientific ou dire que c'est comme dans Alglave 2022 -->

![(\#fig:SchemeRealloc) Schematic representation of the reallocation process. The biomass field (the background field) depends on a covariate and a spatial random effect. The covariate is the $x$ axis. It has a positive effect on biomass values (i.e. biomass is higher on the right of the grid than on the left). The spatial random effect conduct to a hotspot on the bottom-right of the latent field. The study domain is considered as a statistical rectangle (grey square). Fishermen sample catches in areas of poor biomass where the covariate is relatively low (blue points) and in areas of higher biomass where the covariate is higher and eventually in the hotspot of biomass (orange and red points). These catches belong to the same declaration $j$ and are summed to constitute the declaration $D_j=50$. The declaration is declared at the level of the statistical rectangle. From VMS data, we know the fishing positions $x_i$. In standard processing, $D_j$ are then uniformly reallocated over the fishing positions $x_i$. This strongly homogeneizes the catch. In particular, the effect of the habitat is no more evidenced in the reallocated catch $Y_i^*$.](images/realloc.png)

# save(data=gridpolygon_sf_2,file="res/simu_lf.RData")
# 
# plot_1 <- ggplot(gridpolygon_sf_2)+
#   geom_sf(aes(fill=Strue_x),alpha=0.9)+
#   scale_fill_distiller(palette = "Spectral")+
#   theme_void()+theme(legend.position = "none")
# 
# ggsave("images/simu_lf.png",width = 7.5,height = 7.5,dpi = 500)

@@

* the original approach from  @alglave_combining_2022 when the locations of  commercial catch (and eventually the scientific catch) are precisely geolocalized. This ideal situation, with no actual application,  will be named 'Spatial Model' and used as a reference for the comparison between the two alternatives.
* the original model fitted with  commercial reallocated individual catch (and potentially few precisely geolocalized scientific data  as done in  @alglave_combining_2022. This approach will be refereed as "Reallocated Model".
* the alternative approach detailed before where the biomass model is fitted  using commercial catch declaration at a corase spatial level and potentially few precisely geolocalized scientific data. This approach is named "Declaration Model".

@@

In these two sets of simulation studies, there is a unique covariate,  modeled as a continuous GRF that we suppose known at each point of the grid (when present the spatial random component is also a GRF but its values are not observed). 
XXXX AJouter les paramètres du chmaps spatial 
The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$. Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data. All parameterizations are detailed in the Table 1.

@@

We compare the performance of the Spatial Model (the gold standard), the Reallocated Model and the Declaration Model  model configurations in regards to 2 metrics:

@@

@@

df <- data.frame(v1=c("$\\mu$","$\\beta_S$","$\\text{Range}$","$\\text{Marginal variance}$","$\\xi_{com}$","$\\sigma_{com}$","$k_{com}$","$\\xi_{sci}$","$\\sigma_{sci}$"),
                 v2=c("2","2"," -- "," -- ","-1","1"," -- "," -- "," -- "),
                 v3=c("2","2","1.5","0.5","-1","1","1","0","0.8"))

@@

**Single-square simulations**
XXX Infor sur le nombre de simus ???

@@

Furthermore, the number of fishing zones within the statistical rectangle associated with a declaration might also affect the performance of the different approaches. We expect that the reallocation process will be less problematic when all the individual catch are spatially close. This situation corresponds to a declaration associated with only one fishing zone. 
<!-- closed    samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle). Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within the same statistical rectangles. Consequently, a declaration can mix catches that have been realized in zones of high biomass and in zones of low biomass. For instance, a declaration can aggregate data from several fishing operations that occured the same day within the same statistical rectangle, but that were realized on 2 distinct types of fishing grounds (and then on two distinct habitats). Reallocating uniformly the declaration will strongly homogeneise the actual catch, will mix up information from to different grounds and then may lead to a strong loss of accuracy in model outputs. -->
The accuracy of the Reallocated Model outputs is expected to decrease when the number of fishing zones increases.  To assess the effect of such process, we simulated the fishing locations associated with a declaration declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.

XXXX Un petit schéma pour expliquer le  processus de similuation et mettre des images sur les mots de fishing zone, fishing points

@@

```
The resultats are presented in Figures \@ref(fig:ParBiasSingle) and \@ref(fig:ParBiasSingle).

>>>>

@@

Reconciling spatial (and eventually temporal) scales properly when different set of observations do not have the same resolution is a well known issue in geography, ecology, agriculture, geology and statistics [@gotway2002combining]. Inferring values of a variable at points or blocks different from those that have been observed is commonly refered as the *change of support problem* (COS) in the statistical literature [@gotway2002combining,@gelfand2010misaligned]. COS issues can be classified within 2 categories - that are not necessarily mutually exclusive - [@pacifici2019resolving]:

@@

- (2) 'modifiable areal unit' problem (MAUP) when data are aggregated over increasingly larger geographic scales (e.g. collected at point level but regrouped/declared at coarse level - see for instance @gilbert2021integrating). Our case of application clearly refers to MAUP. Note that within this category, another category exist and is commonly refered as the 'ecological fallacy'. This occures when the individual response to a covariate differs from the response estimated from grouping individuals. This means that conclusions from a fine-resolution analysis can differ from an analysis at a coarser scale based on the aggregation of the fine-resolution data.

@@

@@

In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen. This section provides a brief overview of the key ingredient of this model and raises the main concerns on the spatialization of the commercial catch data.

Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ a spatial random field which represents the biomass for a species of interest.  $\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$ and $\bm{\Gamma}=(\Gamma(x),x\in\mathcal{D})$ a field of covariate.  As the catch are zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ with a mixture of a Dirac mass at $0$ and a Log Normal distribution, the proportion of the mixture being defined conditionally on the random field $\bm{S}$.

\begin{equation}
Y_i | S(x_i)  \overset{ind}{\sim} \My{p_i}{\mu_i}{\sigma^2},
\end{equation}

with $p_i \defeq \exp(-e^\xi S(x_{i}))$ the proportion of the mixture, $\mu_i\defeq \frac{S(x_i)}{1-p_i}$ the expected catch when positive and $\sigma^2$ a transformation of its variance. This parametrization (see SMXXXX) of the mixture model corresponds to:
\begin{align}
\P\left (Y_i =0 | S(x_i) \right) & = p_i=  \exp(-e^\xi S(x_{i}) ) \\  \nonumber
\E\left ( Y_i \vert Y_i >0 , S(x_i)\right)  & =  \mu_i = \frac{S(x_i)}{1-p_i},\\  \nonumber
\V ar( Y_i\vert Y_i >0, S(x_i) )& = \mu_i^2 (e^{\sigma^2}-1), \cr \nonumber
\end{align}

@@

@@

$$D=\sum_{i|x_i \in \mathcal{A}_D}{Y_i}$$

Since the original model proposed in @alglave_combining_2022 is defined at the individual catch scale, those catch are derived from the logbooks data combined with the VMS data. This is classically done in fisheries science by using a uniform reallocation process [@hintzen_vmstools_2012,murrayEffectivenessUsingCPUE2013].
This process consists in counting, for a given vessel, on a given day, the number $m_k$ of fishing points in $\mathcal{A}_{D_k}$ associated with declaration $D_k$ and define for each $x_i\in\mathcal{A}_{D_k},$ the associated reallocated individual catch $\Yr_i\defeq D_k/m_k$. As a direct consequence, the so reconstructed individual catch are similar to quantity widely used in fisheries science named Catch Per Unit of Effort (CPUE), effort being measured in hours. As noted by @alglave_combining_2022, this process has several drawbacks.
First, as a consequence of the reallocation process, the reconstructed individual catches tend to exhibit smoother pattern than the original catches. 
Second, the actual number of data is the total number of catch declarations while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations.
From a statistical point of view, this overestismation of the number of informative data tends to produce excessively narrow confidence intervals.

@@

<!-- In standard processing, $D_k$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_k/m_k$. This is what we call uniform (or proportional) reallocation. In this case, inference of species distribution is directly computed through reallocated $Y^*_i$ assuming these are the exact punctual observations. This strongly simplifies the actual process of observation and have most likely repercussion on model performance. -->

<!-- To overcome such limitations, an alternative is to consider that only the catch declarations $D_k$ are observed while $Y_i$ are not (they are latent variables). Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->

To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_k$ instead of the reconstructed individual catch $\Yr_i$. As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_k$ which consists of a sum of $m_k$ random variables following a mixture distribution. This distribution has no known analytical form. However, the catch declaration data also exhibit some zero inflation and a long tail repartition of the values, thus as for $Y_i$ a mixture model is a good candidate to model $D_k$. We define then the distribution of $D_k$ through the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original proposition:

We somehow approximate the sum of the original mixture distributions by a mixture distribution of the same kind and we define the different key quantities, i.e. the mixture proportion, the expected positive catch declaration and its variance in order to match the original proposition:

$$D_k | \mathcal{P}_k,\bm{S} \sim \Md{p^D_k}{\mu^D_k}{\sigma^{2,D}}$$

with $\mathcal{P}_k=(1,...,i, ...,m_{k})$, the list of the fishing positions $(x_1,...,x_i, ...,x_{m_k})$ associated to declaration $D_k=\sum_{i \in \mathcal{P}_k}{Y_i}$, $\mu^D_k$ the expected positive biomass, $p_k^D$ the proportion of the mixture and $\sigma^{2,D}$ the variance parameter.

<!-- that only the catch declarations $D_k$ are observed while $Y_i$ are not (they are latent variables). Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->

<!-- As $D_k=\sum_{i \in \mathcal{P}_k}{Y_{i}}$, it is possible to relate $D_k$, $Y_i$ and $S_{\mathcal{P}_k}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$. Our approach is to match the moments of $D_k$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_k$. This is what we call model-based reallocation. This model will be referred as the Declaration Model. In contrast, the model fitted on reallocated observations $Y^*_i$ will be referred as the $Y^*_i$ model. -->

<!-- In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_k$ instead of $Y_i \vert x_i, S(x_i)$ and $D_k | \mathcal{P}_k,S_{\mathcal{P}_k}$. -->

@@

In order to define relate the individual observation level $Y$ and the catch declaration level $D$, we choose to match the key quantities of the two distributions.

(1) As the $Y_1, \ldots, Y_{m_k}$ are independent conditionally on $\bm{S}$,  the probability to obtain a zero-declaration $\P(D_k = 0\vert \bm{S}, \mathcal{P}_k )$ is obtained by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $i \in \mathcal{P}_k$.

\begin{align*}
\P(D_k = 0\vert \bm{S}, \mathcal{P}_k)  & = \P_{i \in \mathcal{P}_k} P(Y_{i} = 0\vert \bm{S}, x_i),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i \in \mathcal{P}_k} e^{\xi}. S(x_{i})\right \rbrace} = p_k^D
\end{align*}

@@

$$\E(D_k \vert D_k > 0, \bm{S}, \mathcal{P}_k)= \frac{\sum_{i \in \mathcal{P}_k} \E(Y_i \vert S(x_i)) }{1-p_k^D} = \frac{\sum_{i \in \mathcal{P}_k} S(x_{i})}{1-p_k^D},$$
$$\V ar(D_k \vert D_k > 0) = \frac{\sum_{i \in \mathcal{P}_k} \V ar(Y_{i})}{1-p_k^D} - \frac{\pi_k}{(1-p_k^D)^2}E(D_k)^2,$$
$$\mbox{with } \V ar(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i})).$$

Knowing the moment of $D_k$ (that only depends on the distribution of $Y_i$), it is required then to define a probability distribution for $D_k|D_k>0$ that is adapted to the underlying individual observations and express $\P(D_k|D_k>0)$ as a function of $\E(D_k \vert D_k > 0, \bm{S}, \mathcal{P}_k)$ and $\V ar(D_k \vert D_k > 0)$. We propose to use the same family of distribution for $D_k|D_k>0$ than the one used for the individual catch $Y_i|Y_i>0$ i.e. a lognormal distribution. This is an approximation that we discuss hereafter.

By defining the observation process at the declaration level, we expect to avoid some of the drawbacks of the estimation based on reallocated individual catch data $\Yr$.

Finally, note that the scientific data is considered known at each fishing locations and thus they can be simply modelled as standard individual data $Y_i$.

The inference is based on  maximum likelihood approach with two simplifications. We use the Stochastic Partial Differential Equations (SPDE) approach to represent the spatial field [@lindgren_explicit_2011] and we use Laplace approximation to compute the likelihood. The stochastic random field is also approximated by a piecewise constant process defined on a fine grid. The optimization of the likelihood relies Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].

![(\#fig:SchemeRealloc) Schematic representation of the reallocation process. The biomass field (the background field) depends on a covariate and a spatial random effect. The covariate is the $x$ axis. It has a positive effect on biomass values (i.e. biomass is higher on the right of the grid than on the left). The spatial random effect conduct to a hotspot on the bottom-right of the latent field. The study domain is considered as a statistical rectangle (grey square). Fishermen sample catches in areas of poor biomass where the covariate is relatively low (blue points) and in areas of higher biomass where the covariate is higher and eventually in the hotspot of biomass (orange and red points). These catches belong to the same declaration $j$ and are summed to constitute the declaration $D_k=50$. The declaration is declared at the level of the statistical rectangle. From VMS data, we know the fishing positions $x_i$. In standard processing, $D_k$ are then uniformly reallocated over the fishing positions $x_i$. This strongly homogeneizes the catch. In particular, the effect of the habitat is no more evidenced in the reallocated catch $Y_i^r$.](images/realloc.png)

@@

* the original approach from @alglave_combining_2022 when the locations of commercial catch (and eventually the scientific catch) are precisely geolocalized. This ideal situation, with no actual application, will be named 'Spatial Model' and used as a reference for the comparison between the two alternatives.
* the original model fitted with commercial reallocated individual catch (and potentially few precisely geolocalized scientific data) as done in @alglave_combining_2022. This approach will be refered as "Reallocated Model".
* the alternative approach detailed before where the biomass model is fitted  using commercial catch declaration at a coarse spatial level and potentially few precisely geolocalized scientific data. This approach is named "Declaration Model".

@@

In these two sets of simulation studies, there is a unique covariate,  modeled as a continuous GRF that we suppose known at each point of the grid (when present the spatial random component is also a GRF but its values are not observed).

The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$. Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data. All parameterizations are detailed in the Table 1.

@@

We compare the performance of the Spatial Model (the gold standard), the Reallocated Model and the Declaration Model configurations in regards to 2 metrics:

@@

To get enough replicates, we run the simulations 100 time.

@@

df <- data.frame(v1=c("$\\mu$","$\\beta_S$","$\\text{Range } (\\delta)$","$\\text{Marginal variance }  (\\delta)$","$\\text{Range } (\\Gamma)$","$\\text{Marginal variance }  (\\Gamma)$","$\\xi_{com}$","$\\sigma_{com}$","$k_{com}$","$\\xi_{sci}$","$\\sigma_{sci}$"),
                 v2=c("2","2"," -- "," -- ","10 (cells)","0.5","-1","1"," -- "," -- "," -- "),
                 v3=c("2","2","0.6 ($\\approx$ 50 km)","1","1.5 ($\\approx$ 120 km)","0.5","-1","1","1","0","0.8"))

@@

**Single-square simulations**

@@

Furthermore, the number of fishing zones within the statistical rectangle associated with a declaration might also affect the performance of the different approaches. We expect that the reallocation process will be less problematic when all the individual catch are spatially close. This situation corresponds to a declaration associated with only one fishing zone. 
<!-- closed    samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle). Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within the same statistical rectangles. Consequently, a declaration can mix catches that have been realized in zones of high biomass and in zones of low biomass. For instance, a declaration can aggregate data from several fishing operations that occured the same day within the same statistical rectangle, but that were realized on 2 distinct types of fishing grounds (and then on two distinct habitats). Reallocating uniformly the declaration will strongly homogeneise the actual catch, will mix up information from to different grounds and then may lead to a strong loss of accuracy in model outputs. -->
The accuracy of the Reallocated Model outputs is expected to decrease when the number of fishing zones increases.  To assess the effect of such process, we simulated the fishing locations associated with a declaration declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones (Figure \@ref(fig:SimuSevZone)).

## 1 zone
load(file="res/fish_1zones.RData")

centres <- fish_zones_res$centres
peche_com_old <- fish_zones_res$peche_com_old
gridpolygon_sf <- fish_zones_res$gridpolygon_sf

centres_sf <- st_as_sf(centres,coords = c("x","y"))
buffer_sf <- st_buffer(centres_sf,dist = 3,endCapStyle = 'SQUARE')

gridpolygon_sf_2 <- gridpolygon_sf[st_intersects(gridpolygon_sf,buffer_sf) %>% lengths > 0,] %>%
  mutate(centr = 1) %>%
  dplyr::select(-layer)

one_zones_plot <- ggplot()+
  geom_point(data=centres,aes(x=x, y=y, col=as.factor(boats)),
             shape = 4,size=4)+
  geom_point(data=peche_com_old,aes(x=x, y=y, col=as.factor(boats)), size=4)+
  geom_sf(data = gridpolygon_sf,alpha=0)+
  geom_sf(data = gridpolygon_sf_2,alpha=0.2,fill="red")+
  theme_void()+
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))+
  ggtitle("One zone")

## 3 zones
load(file="res/fish_3zones.RData")

centres <- fish_zones_res$centres
peche_com_old <- fish_zones_res$peche_com_old
gridpolygon_sf <- fish_zones_res$gridpolygon_sf

centres_sf <- st_as_sf(centres,coords = c("x","y"))
buffer_sf <- st_buffer(centres_sf,dist = 3,endCapStyle = 'SQUARE')

gridpolygon_sf_2 <- gridpolygon_sf[st_intersects(gridpolygon_sf,buffer_sf) %>% lengths > 0,] %>%
  mutate(centr = 1) %>%
  dplyr::select(-layer)

three_zones_plot <- ggplot()+
  geom_point(data=centres,aes(x=x, y=y, col=as.factor(boats)),
             shape = 4,size=4)+
  geom_point(data=peche_com_old,aes(x=x, y=y, col=as.factor(boats)), size=4)+
  geom_sf(data = gridpolygon_sf,alpha=0)+
  geom_sf(data = gridpolygon_sf_2,alpha=0.2,fill="red")+
  theme_void()+
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))+
  ggtitle("Three zones")

## 5 zones
load(file="res/fish_5zones.RData")

centres <- fish_zones_res$centres
peche_com_old <- fish_zones_res$peche_com_old
gridpolygon_sf <- fish_zones_res$gridpolygon_sf

centres_sf <- st_as_sf(centres,coords = c("x","y"))
buffer_sf <- st_buffer(centres_sf,dist = 3,endCapStyle = 'SQUARE')

gridpolygon_sf_2 <- gridpolygon_sf[st_intersects(gridpolygon_sf,buffer_sf) %>% lengths > 0,] %>%
  mutate(centr = 1) %>%
  dplyr::select(-layer)

five_zones_plot <- ggplot()+
  geom_point(data=centres,aes(x=x, y=y, col=as.factor(boats)),
             shape = 4,size=4)+
  geom_point(data=peche_com_old,aes(x=x, y=y, col=as.factor(boats)), size=4)+
  geom_sf(data = gridpolygon_sf,alpha=0)+
  geom_sf(data = gridpolygon_sf_2,alpha=0.2,fill="red")+
  theme_void()+
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))+
  ggtitle("Five zones")

sev_zones_plot <- plot_grid(one_zones_plot,
                            three_zones_plot,
                            five_zones_plot,nrow = 1)

ggsave("images/sev_zones_plot.png",width = 9, height = 3)

![(\#fig:SimuSevZone) Simulations of 10 fishing points within 1, 3 and 5 fishing zones. The full grid correspond to a statistical rectangle. Cross are the centroid of the fishing zones. A declaration declared at the level of the statistical rectangle would be uniformly reallocated over these fishing points.](images/sev_zones_plot.png)

@@

The resultats are presented in Figures \@ref(fig:PerfMetricSingle) and \@ref(fig:ParBiasSingle).
